---
title: "Untitled"
author: '109090046'
date: "2023-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

```{r}
myplot = function(m, s, n = 500)
{
  
  r = rlnorm(n,m,s) 
  q = quantile(r,seq(0,1,0.01))
  p = dlnorm(q,m,s) 
  
  
  hist(r, freq = F, main = "", xlab = "", xlim = c(0,5))
  lines(density(r), main = "", xlab = "")
  lines(x = q, y = p, col = "red")
}

par(mfrow = c(2,2))

myplot(m = 0, s = 0.2)
myplot(m = 0, s = 0.3)
myplot(m = 1, s = 0.2)
myplot(m = 1, s = 0.3)


```

`m`: The mean of the log-normal distribution.
`s`: The standard deviation of the log-normal distribution.
`n`: The number of observations to generate. It has a default value of 500.

`par(mfrow = c(2,2))` is used to divide the plotting region into a 2x2 grid, so that four plots can be displayed in the same window.

## 2
**Ans**: Two line plots and one histogram.

## 3
**Ans**: xlim=c(0,5) is to set the histograms x-axis to show from value 0 to 5.

## 4
**Ans**: To let the total area in the histogram to be 1, which is set each bins value by percentages.

## 5

```{r}
myplot = function(m, s, n = 500)
{
  
  r = rnorm(n,m,s) 
  q = quantile(r,seq(0,1,0.01))
  p = dlnorm(q,m,s) 
  
  
  hist(r, freq = F, main = "", xlab = "", xlim = c(0,5))
  lines(density(r), main = "", xlab = "")
  lines(x = q, y = p, col = "red")
}

par(mfrow = c(2,2))

myplot(m = 2, s = 0.2, n=1000)
myplot(m = 2, s = 0.4, n=1000)
myplot(m = 3, s = 0.2, n=1000)
myplot(m = 3, s = 0.4, n=1000)

```

## 6

```{r}
callprice <- function(xt = 10, K = 10, t = 0, T = 1, R = 0.03, eta = 0.02, sigma = 0.4) {
  
  d1 <- (log(xt/K) + (R - eta + 0.5*sigma^2)*(T-t)) / (sigma*sqrt(T-t))
  d2 <- d1 - sigma*sqrt(T-t)
  
  N1 <- pnorm(d1)
  N2 <- pnorm(d2)
  
  call_price <- xt*exp(-eta*(T-t))*N1 - K*exp(-R*(T-t))*N2
  
  return(call_price)
  
}

ct = function(x,K)
{
  callprice(sigma = x, K = K)
}

curve(ct(x,K = 10), from = 0, to = 20, n = 200, lwd=2)
curve(ct(x,K = 5), from = 0, to = 20, n = 200, add = T, col = "green", lwd=2)
curve(ct(x,K = 15), from = 0, to = 20, n = 200, add = T, col = "red", lwd=2)
```

3 fingers
