---
title: "quiz3"
author: '109090046'
date: "2023-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      comment = "", 
                      dev.args = list(bg = 'lavender',                                       pointsize = 11))

library(tidyverse)
library(nycflights13)
library(magrittr)
library(kableExtra)
ptable = function(df,digits = getOption("digits"),size = 14){
  df %>% knitr::kable(digits = digits) %>% 
    kable_classic(lightable_options = c("striped", "hover", "condensed"),
                  fixed_thead = list(enabled = T, 
                                     background = "lavender"),
                  font_size = size, full_width = F,
                  html_font = "helvetica")
}
```

## 1

```{r}
flights %>% transmute(across(ends_with("time")& -air_time,
                             ~60*as.integer(str_sub(.x,1,-3))+as.integer(str_sub(.x,-2,-1)),
                             .names = "new_{.col}"))
```

new_dep_time: The departure time of the flight in minutes since midnight.
new_arr_time: The arrival time of the flight in minutes since midnight.
new_sched_dep_time: The scheduled departure time of the flight in minutes since midnight.
new_sched_arr_time: The scheduled arrival time of the flight in minutes since midnight.

## 2

You can also embed plots, for example:

```{r}
library(gapminder)

dd <- c('red', 'gray', 'yellow', 'cyan', 'white')
names(dd) <- c('Africa', 'Americas', 'Asia', 'Europe', 'Oceania')

years <- gapminder %>% filter(year %in% c(1952, 1977, 2002))

par(mfrow = c(1, 3), mar = c(4.5, 4.5, 1.2, 2),
    bg = "black", fg = "white", 
    col.lab = "white", col.main = "white", col.axis = "white")

for (yr in c(1952, 1977, 2002)) {
    years %>% 
    filter(year == yr) %$% 
    plot(y = lifeExp, x = log10(gdpPercap), xlim = c(3, 5), ylim = c(35, 85),
         xlab = "log10(GDP per Capita)", ylab = "Life Expectancy",
         pch = 16, main = paste0("Year = ", yr),
         col = as.vector(recode(continent, !!!dd)))
}

par(bg = "black", fg = "white", 
    col.lab = "white", col.main = "white", col.axis = "white")

legend("bottomright", inset = c(0.05, 0.1),
       pch = 16, ncol = 1, 
       col = dd, y.intersp = 0.8,
       title = "Continent",
       legend = names(dd))

```

This code produces a similar plot as before, but with a logarithmic transformation applied to the x-axis using the log10() function. The xlim parameter is set to restrict the x-axis to a range of 3 (corresponding to a GDP per capita of 1000) to 5 (corresponding to a GDP per capita of 100000).