---
title: "Financial Big Data HW2"
author: '109090046呂庭羽'
date: "2023-03-01"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Reading dataset
Read the csv file into a data.frame by the name market.

```{r data}
library(readr)
market <- read_csv("D:/下載/marketing_data.csv", show_col_types = FALSE)
```

---

### Question 1
Use both `str()` and `skim()` to check the structure of 'market'.

```{r q1}
#Use `str()`
str(market)

#Use `skim()`
library(skimr)
skim(market)
```

---

### Question 2
List all name of variables in 'market'.

```{r q2}
names(market)
```

---

### Question 3
Find out the number of rows that contain missing values.

```{r q3}
sum(!complete.cases(market))
```

---

### Question 4
Find out the number of missing values in each column.

```{r q4}
colSums(is.na(market))

library(skimr)
skim(market)
```

---

### Question 5
Use the `describe()` from 'psych' to check 'market' and explain how it handles character variables. Also, use `summary()` to see how it handles character variables.

```{r q5}
library(psych)
describe(market)

summary(market)
```
`describe()` shows the mean, standard deviation, median, trimmed(mean without outliers), median absolute deviation, range, skewness, kurtosis and standard error with data.frame, while `summary()` shows quartile.

---

### Question 6
Use `table()` on 'Education', 'Martial_Status' and 'Country'.
And on the pair of 'Education' and 'Martial_Status'.
Redo them with `%$%`.

```{r q6}
table(market$Education)
table(market$Marital_Status)
table(market$Country)

table(market$Education, market$Marital_Status)

library(magrittr)
table(market%$%Education)
table(market%$%Marital_Status)
table(market%$%Country)

table(market%$%Education, market%$%Marital_Status)
```

`table()` creates frequency table of each character variable.
And, `table()` on the pair of character variables shows the frequency table for two character variables mutually.

---

### Question 7

```{r q7}
income <- market$Income
head(income)

market$Income <- as.numeric(gsub("[^0-9.]", "", market$Income))
head(market$Income)
```
`Income` in csv are formatted as currency which always has a `$` in front of every numeric data, so after we imported it into R they will be imported as character strings.

The `gsub()` function removes any non-numeric characters (including the currency symbol) from the strings, and the `as.numeric()` function converts the cleaned-up strings to numeric values.Therefore, the values should be numeric now.

---

### Question 8

```{r q8}
library(skimr)
Nunique = apply(market,2,n_unique)
col = colnames(market)[Nunique == 2]
tbl = apply(market[,col],2,table)
ratio = apply(market[,col],2,mean)
```

```{r Nunique}
Nunique
```
**Nunique**: The number of unique value in each character variable.
```{r col}
col
```
**col**: The name of each character variable which contains only 2 unique values.
```{r tbl}
tbl
```
**tbl**: To count the number of 0 and 1 in each character variable in `col` make them a data.frame.
```{r ratio}
ratio
```
**ratio**: The ratio of 1 in each character variable in `col`.

---

### Question 9

```{r q9-1}
myf = function(var,byvar)
{
    by = unique(byvar)
    mout = NULL
    for (i in by) {
        s = subset(var,byvar == i)
        m = colMeans(s)
        mout = cbind(mout,m)
    }
    colnames(mout) = by
    mout
}
```

```{r q9-2}
varlist = c("MntWines","MntSweetProducts",
            "MntMeatProducts","MntFishProducts",
            "MntGoldProds")
colspec = c(4,2,1,5,8)
tbl = myf(subset(market,,varlist),market$Education)
```

```{r q9-3}
tbl
barplot(tbl, beside = T, col = colspec)
legend("topright", inset = c(0,0), legend = varlist, 
       col = colspec, pch = 16, cex = 1.1)
```

```{r q9-4}
agegrp = ifelse(market$Year_Birth < 1960,"<1960",
                ifelse(market$Year_Birth < 1970,"[1960,1970)",
                       ifelse(market$Year_Birth < 1980,"[1970,1980)",">=1980")))
```

**agegrp** shows how when every customer is born and seperates them into 4 groups, which are people who are born before 1960, after 1980, between 1960 and 1970, and between 1970 and 1980.

```{r agegrp}
agegrp

table(agegrp)
```

```{r q9-5}
varlist2 <- c("Kidhome", "Teenhome")
tbl2 = myf(subset(market,,varlist2),agegrp)
```

```{r tbl2}
tbl2
```
**tbl2**(just to distinguish from tbl) shows people from each age group have the average number of kids or teenagers in home.

```{r tbl2_plot}
tbl2
barplot(tbl2, beside = T, col = colspec[0:2])
legend("topright", inset = c(0,0), legend = varlist2, 
       col = colspec[0:2], pch = 16, cex = 1.1)
```